---
title: "Holothruia atra Anlaysis"
author: "Nikko"
date: "7/11/2021"
output:
  pdf_document: default
  html_document: default
---

\#\#Set up

```{r}
library(knitr)
library(rentrez)
library(pegas)
library(tidyverse)
library(geomedb)
library(strataG)
```

# Query metadata

Just looked through general longitude and latitudes to get majority of the data

```{r}
holothuria_metadata <- queryMetadata(
  entity = "Sample",
  query = "genus = Holothuria AND specificEpithet = atra AND _exists_:sequence",
  select = "Event")
View(holothuria_metadata$Event)
View(holothuria_metadata$Sample)


```

## Subset the metadata

```{r}
holothuria_event_info <-holothuria_metadata$Event[,c("locality", "country", "decimalLatitude", "decimalLongitude", "eventID")]
kable(holothuria_event_info)
```

# Lumping and cleaning locality names

```{r}
#Lump Halmahera samples together
holothuria_metadata$Event$locality[grep("Halmahera",holothuria_metadata$Event$locality)] <- "Halmahera"
holothuria_metadata$Event$locality <- str_remove_all(holothuria_metadata$Event$locality,'\\"')   %>% str_remove_all("[(,)]")

#join Sample and Event tables
holothuria_flat <- left_join(holothuria_metadata$Sample, holothuria_metadata$Event, by = "eventID")

#remove samples from populations with n < 5
holothuria_flat <- holothuria_flat %>% filter(locality != "Gardner") %>% filter(locality != "New Caledonia") %>% 
            filter(locality != "Nusa Penida") %>% filter(locality != "Papua New Guinea") %>% 
            filter(locality != "Rakata Krakatau Besar") %>% filter(locality != "Solomon Islands") %>% 
            filter(locality != "Tile-tile Selayar") %>% filter(locality != "Yop T. Cenderwasih")

```

# Query FASTA

```{r}
listLoci()
```

```{r}
holothuria_CO1 <- querySanger(query = "genus = Holothuria AND specificEpithet = atra AND _exists_:sequence", 
                              locus = "CO1")
holothuria_CO1
```

\#\#Genetic data

```{r}
#Overwrite the names from the FASTA to have shorter values
names(holothuria_CO1) <- str_replace(string=names(holothuria_CO1),
                                     pattern=".+\\[tissueID = (.+?)\\].+", replacement = "\\1")
```

# Sequence alignment

```{r}
#Align the data to ensure that each site is homologous
holothuria_CO1 <- muscle(holothuria_CO1)

image.DNAbin(holothuria_CO1)
```

150 to 200 looks interesting...

```{r}
image.DNAbin(holothuria_CO1[1:20, 150:200])
```

Let's trim the missing parts of the Skillings sequences

```{r}
image.DNAbin(holothuria_CO1[,60:482])

holothuria_CO1 <- holothuria_CO1[,60:482]

```

Put the metadata in the same order as the genetic data, and only keep the sequences that are still found in the metadata after we removed the ones from pops with n \< 5

```{r}
holothuria_flat <- holothuria_flat[order(holothuria_flat$materialSampleID),]
holothuria_CO1 <- holothuria_CO1[which(rownames(holothuria_CO1) %in% holothuria_flat$materialSampleID),]
holothuria_CO1 <- holothuria_CO1[order(rownames(holothuria_CO1)),]
```

## Creating a haplotype network

```{r}
#Make a list of populations that were sampled with locality and country names
holothuria_pop <- holothuria_flat$locality
#write out population map for editing
write.csv(holothuria_pop,"Holothuria_atra_Population_Map.csv", quote=F,row.names = F)
#read in edited population map
#holothuria_pop_maps <- read.csv("Holothuria_atra_Population_Map.csv")
#holothuria_pop <- holothuria_pop_maps$regions

a <- dist.dna(holothuria_CO1)
b <- pegas::haplotype(holothuria_CO1)
b <- sort(b, what="label")

#Create the network
holothuria_net <- pegas::haploNet(b)

#Make a table of which haplotypes occur in which populations
c <-stack(setNames(attr(b, "index"), rownames(b)))
c <-c[order(c$values),]
holo_ind.hap<-table(hap=c$ind, holothuria_pop=holothuria_pop)



#Skip distinguishing oceans for now?
plot(holothuria_net,size=attr(holothuria_net, "freq"), scale.ratio=0.5, pie=holo_ind.hap, legend=F, labels=F, threshold=0, show.mutation=2, fast=T)
```

# Prepping for PopArt

```{r}
hapseqs <- b[]
write.nexus.data(hapseqs, file = "holothuria_haplotypes4.nex", interleaved = F)
write.csv(holo_ind.hap, file = "holothuria_map4.csv", quote = F)
#Do I need to do below? It classifies the oceans
#write.csv(holo_ind.hap2, file="holothuria_map_IP.csv", quote = F)
holothuria_event_info <- unique(holothuria_flat[,c("locality",
                                                   "decimalLatitude","decimalLongitude")])
write.csv(holothuria_event_info, file = "holothuria_event_metadata.csv", quote = F)
```

# Diversity statistics

## Haplotype diversity

```{r}
#Create a copy of dataset
holothuria_CO1_copy <- holothuria_CO1
rownames(holothuria_CO1_copy) <- holothuria_pop
stratastat<-function(x,pop=pop, fun=nuc.div){
  stats<-NULL
  for(p in unique(pop)){
    print(p)
    stat<-fun(x[grep(p,rownames(x)),])
    stats <- c(stats,stat)
  }
  
  names(stats)<-unique(holothuria_pop)
  return(stats)
}

holothuria_hapdivs <- stratastat(holothuria_CO1_copy, pop = holothuria_pop, fun = hap.div)

holothuria_hapdivs
```

# Nucleotide diversity

```{r}
holothuria_nucdivs <- stratastat(holothuria_CO1_copy, pop = holothuria_pop, fun = nuc.div)
holothuria_nucdivs

divs <- cbind(holothuria_nucdivs, holothuria_hapdivs)

write.csv(divs,"Hap_Nuc_Diversities.csv")

```

# Convert to gtypes

```{r}
holothuria_z<-sequence2gtypes(holothuria_CO1, strata=holothuria_pop)
holothuria_z<-labelHaplotypes(holothuria_z)
holothuria_z
```

# PhiST table

```{r}
pairwise <- pairwiseTest(holothuria_z)

pairwise_phi <- pairwiseMatrix(pairwise, stat="PHIst")

#PhiST in lower triangle, p-value in upper triangle
kable(pairwise_phi, digits=3)
```

# Saving as table

```{r}
library(kableExtra)

kable(pairwise_phi, digits = 3, "html", align=c("c", "c", "c", "c","c", "c", "c", "c","c", "c", "c", "c","c", "c", "c", "c","c", "c", "c", "c","c", "c", "c", "c", "c", "c","c", "c", "c", "c")) %>%
  kable_styling("striped") %>%
  as_image()
```

```{r}
phiST<-pairwise_phi
phiST[upper.tri(phiST)] = t(phiST)[upper.tri(phiST)]
heatmap(phiST, margins = c(11,11))
```

```{r}
divs
```

```{r}
holothuria_hapdivs
```

# Organized table on excel holothuria_table

{r}
colnames(holothuria_table) \<- c(" ", "\# Samples", "\# Haplotypes", "Nucleotide Diversity", "Haplotype Diversity")
diversities\<-list(holothuria_table)
diversities
kable(diversities, digits=3)


kable(diversities, digits = 3, "html") %\>%
kable_styling("striped") %\>%
as_image()

\#Will not save!

```{r}
#Doing it manually
stratum <-c("Alam Kotok Pulau Seribu", "Bunaken-Liang Beach", "Eastern Australia", "French Frigate", "Gili Trawagan Lombok", "Halmahera", "Hilo", "Ishigaki", "Jarvis", "Johnston", "Karang Congkak", "Karimunjawa-Pulau Tengah", "Kauai", "Kingman", "Kodingarengkeke Makassar", "Kona", "Kukusan Waiciciu Flores", "Kure", "Kwajalein", "Lahewa Nias", "Laysan", "Majuro", "Mayu Island Maluku Sea", "Midway", "Niihau", "Oahu", "Ogasawara","Palmyra", "Pearl & Hermes", "Pulau Weh Aceh")
individuals <-c("21", "21", "11", "28", "5", "40", "9", "24", "23", "26", "17", "18", "30", "27", "10", "21", "19", "23", "9", "7", "12", "13", "13", "35", "5", "24", "20", "16", "37", "20")
haplotypes <-c("3", "5", "3", "10", "2","6", "4", "3", "5", "7", "3", "5", "8", "10", "2", "10", "3", "8", "4", "3", "8", "4", "1", "14", "4", "7", "4", "4", "10", "4")
nd <-c("0.007", "0.012", "0.004", "0.008", "0.001", "0.007", "0.004", "0.003", "0.013", "0.013", "0.011", "0.011", "0.003", "0.014", "0.008", "0.008", "0.003", "0.010", "0.013", "0.011", "0.006", "0.012", "0.000", "0.008", "0.007", "0.005", "0.013", "0.016", "0.009", "0.004")
hd <-c("0.624", "0.738", "0.618", "0.881", "0.600", "0.752", "0.806","0.507","0.652", "0.812", "0.640", "0.791", "0.747","0.897","0.356", "0.867", "0.284", "0.854", "0.750", "0.667", "0.894", "0.758","0.000", "0.839", "0.900", "0.790","0.658","0.758","0.788", "0.574")

diversities_table <- data.frame(stratum, individuals, haplotypes, nd, hd)
diversities_table
kable(diversities_table, "html", digits = 3, col.name=c("Stratum", "# Samples", "# Haplotypes", "Nucleotide Diversity", "Haplotype Diveristy")) %>%
  kable_styling("striped") %>%
  as_image()

```
